---
title: "Final Project MGSC 310"
author: "Alejandro Rea"
subtitle: Final Project using Spotify Data to see What Influences Song Skipability
output:
  html_document:
    df_print: paged
  html_notebook: default
---

```{r setup, include=FALSE}

# Please leave this code chunk as is. It makes some slight formatting changes to alter the output to be more aesthetically pleasing. 

library(knitr)

# Change the number in set seed to your own favorite number
set.seed(1818)
options(width=70)
options(scipen=99)


# this sets text outputted in code chunks to small
opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=TRUE, size = "vsmall")  
opts_chunk$set(message = FALSE,                                          
               warning = FALSE,
               # "caching" stores objects in code chunks and only rewrites if you change things
               cache = FALSE,                               
               # automatically downloads dependency files
               autodep = TRUE,
               # 
               cache.comments = FALSE,
               # 
               collapse = TRUE,
               # change fig.width and fig.height to change the code height and width by default
               fig.width = 5.5,  
               fig.height = 4.5,
               fig.align='center')


```

```{r setup-2}

# Always print this out before your assignment
sessionInfo()
getwd()

```


<!-- ### start answering your problem set here -->
<!-- You may export your homework in either html or pdf, with the former usually being easier. 
     To export or compile your Rmd file: click above on 'Knit' then 'Knit to HTML' -->
<!-- Be sure to submit both your .Rmd file and the compiled .html or .pdf file for full credit -->


```{r setup-3}

# load all your libraries in this chunk 
library('tidyverse')
library('dplyr')
library('readr')
library('rsample')
library('vtable')
library('randomForest')
library('xgboost')
library('ROSE')
# note, do not run install.packages() inside a code chunk. install them in the console outside of a code chunk. 

```



## Question 1

1a) Text response to part a. 

```{r}

track_features <- read_csv("datasets/tf_mini.csv")
session_logs <- read_csv("datasets/log_mini.csv")

#renaming track_id to link to track_features dataset
names(session_logs)[names(session_logs) == "track_id_clean"] <- "track_id"


#track and session data together
songs <- merge(track_features, session_logs, by="track_id")

#checking for missing data - no null data
#colSums(is.na(songs))

#omitting any NA value rows just in case
songs_clean <- na.omit(songs)


```

```{r}

st(songs_clean)

```


1b) Response to part b. 

```{r}

#turning our variables into yes or no in a binary fashion. We could just factor if you fancy that but
songs_clean <- songs_clean %>% 
  mutate( skip_3 = ifelse(skip_3 == "TRUE", 1,0),
          not_skipped = ifelse(not_skipped == "TRUE", 1,0))



```

```{r}

#Since we only care if the song will be counted as a stream we made a new attribute that identifies that. A song will be a stream if it is played for 30 seconds, so if the song is briefly played and skipped, played for a while and skipped, or not skipped at all, it will count as a stream for the artist.

songs_clean$counted_stream <- ifelse((songs_clean$skip_3 == 1 | songs_clean$not_skipped == 1), 1, 0)



#THIS DATASET IS IF YOU WANT TO KEEP ALL THE USER BEHAVIOR VARIABLES
songs_clean_complete <- songs_clean %>% 
  mutate(counted_stream = as_factor(counted_stream),
         mode = as_factor(mode),
         time_signature = as_factor(time_signature),
         hist_user_behavior_is_shuffle = as_factor(hist_user_behavior_is_shuffle),
         premium = as_factor(premium),
         context_type = as_factor(context_type),
         hist_user_behavior_reason_start = as_factor(hist_user_behavior_reason_start),
         hist_user_behavior_reason_end = as_factor(hist_user_behavior_reason_end)
         )

#FURTHER REMOVING UNNECESSARY VARIABLES FROM THE DATAFRAME THAT USES SONG AND USER BEHAVIOR
songs_clean_complete <- songs_clean_complete %>% select(-acoustic_vector_0,-acoustic_vector_1,-acoustic_vector_2,-acoustic_vector_3,-acoustic_vector_4, -acoustic_vector_5, -acoustic_vector_6, -acoustic_vector_7, -track_id, -session_id, -date, -skip_1, -skip_2, -skip_3, -not_skipped, -release_year)


#THIS DATASET IS IF YOU WANT TO REMOVE USER BEHAVIOR VARIABLES AND WANT TO KEEP ONLY SONG DATA
songs_clean <- songs_clean %>% select(-track_id, -session_id, -date, -skip_1, -skip_2, -skip_3, -not_skipped, -session_position, -session_length, -context_switch, -no_pause_before_play, -short_pause_before_play, -long_pause_before_play, -hist_user_behavior_n_seekfwd, -hist_user_behavior_n_seekback, -hist_user_behavior_is_shuffle, -hour_of_day, -premium, -context_type, -hist_user_behavior_reason_start, -hist_user_behavior_reason_end, -release_year, -acoustic_vector_0,-acoustic_vector_1,-acoustic_vector_2,-acoustic_vector_3,-acoustic_vector_4, -acoustic_vector_5, -acoustic_vector_6, -acoustic_vector_7)


#FACTORING ALL THE NECESSARY VARIABLES FOR THE SONG ONLY DATA (DOES NOT INCLUDE USER BEHEVAIOR)
songs_clean <- songs_clean %>%
  mutate(counted_stream = as_factor(counted_stream),
        mode = as_factor(mode),
         time_signature = as_factor(time_signature),
        )



```

## undersampling the dataframe that uses song and user data
```{r}

over <- ovun.sample(counted_stream ~ . , data = songs_train, method = "over", N = 247104)$data
table(over$counted_stream)

```

## undersampling the dataframe that only uses song data
```{r}

under <- ovun.sample(counted_stream ~ . , data = songs_clean, method = "under", N = 6292)$data
table(under$counted_stream)
```

## Both Oversampling and Undersampling using both song and user data
```{r}

both <- ovun.sample(counted_stream ~ . , data = songs_train, method = "both", N = 125910)$data
table(both$counted_stream)

```

```{r}

#SPLIT FOR THE DATAFRAME THAT INCLUDES SONGS AND USER BEHAVIOR VARIABLES
songs_split_under <- initial_split(under, prop = 0.75) 

songs_train_under <- training(songs_split_under)
songs_test_under <- testing(songs_split_under)

#SPLIT FOR THE DATAFRAME THAT ONLY USES SONG VARIABLES
songs_split <- initial_split(songs_clean, prop = 0.75) 

songs_train <- training(songs_split)
songs_test <- testing(songs_split)




```


```{r}

print(songs_train_completeDF)

unique(songs_train_completeDF$hist_user_behavior_reason_start)

unique(songs_train_completeDF$hist_user_behavior_reason_end)

```

etc...

##ILLUSTRATES OUR CLASS IMBALANCE
```{r}

class(songs_clean$counted_stream)

# 
# barplot(prop.table(table(under$counted_stream)),
#         col = rainbow(2),
#         main = "Counted Stream Distribution")

prop.table(table(songs_clean$counted_stream))
```





## Question 2

my random forest that uses both user and song data WITH NO RESAMPLING

```{r}

rf_fit_songs <- randomForest(counted_stream ~ .,
                        data = songs_train_completeDF,
                        na.action = na.roughfix,
                        importance = TRUE)

print(rf_fit_completeDF)

```

my random forest that uses ONLY song data WITH NO RESAMPLING

```{r}


rf_fit_songs <- randomForest(counted_stream ~ .,
                        data = songs_train,
                        mtry = 11,
                        ntree = 1500,
                        na.action = na.roughfix,
                        importance = TRUE)

print(rf_fit_songs)

```


random forest using only the music related data
```{r}

rf_fit_only_songs <- randomForest(counted_stream ~ .,
                        data = songs_train,
                        na.action = na.roughfix,
                        importance = TRUE)



```

```{r}

print(rf_fit_only_songs)

```


Random forest that uses user and song data USING OVERSAMPLING
```{r}

rf_fit_overSample <- randomForest(counted_stream ~ .,
                        data = over,
                        na.action = na.roughfix,
                        importance = TRUE)

print(rf_fit_overSample)

```


CROSS VALIDATING
```{r}

rf_modsLab <- list()
oob_errLab <- NULL
test_errLab <- NULL
for(mtry in 1:25){
  rf_fitLab <- randomForest(counted_stream ~ ., 
                         data = songs_train_under,
                         mtry = mtry,
                         na.action = na.roughfix,
                         ntree = 600)
  oob_errLab[mtry] <- rf_fitLab$err.rate[600]
  
  cat(mtry," ")
}
```

```{r}

results_DF <- data.frame(mtry = 1:25, oob_errLab)
ggplot(results_DF, aes(x = mtry, y = oob_errLab)) + geom_point() + theme_minimal()


```

Random forest that uses only song data USING UNDERSAMPLING
```{r}

rf_fit_underSample <- randomForest(counted_stream ~ .,
                        data = songs_train_under,
                        mtry = 11,
                        ntree = 1500,
                        na.action = na.roughfix,
                        importance = TRUE)

print(rf_fit_underSample)

```
Random forest that uses only song data USING UNDERSAMPLING on TEST SET to see OOB vs TRAIN SET OOB
```{r}

rf_fit_underSampleTest <- randomForest(counted_stream ~ .,
                        data = songs_test_under,
                        mtry = 11,
                        ntree = 3000,
                        na.action = na.roughfix,
                        importance = TRUE)

print(rf_fit_underSampleTest)

```

Random forest that uses user and song data USING BOTH OVERSAMPLING AND UNDERSAMPLING
```{r}

rf_fit_both <- randomForest(counted_stream ~ .,
                        data = both,
                        na.action = na.roughfix,
                        importance = TRUE)

print(rf_fit_both)

```

```{r}
print(rf_fit_only_songs)

```

2b) Response to part b. 

```{r}

print(rf_fit_songs)


```

```{r}

table(over$counted_stream)

table(songs_train$counted_stream)

table(songs_train_completeDF$counted_stream)


table(songs_clean_complete$counted_stream)

```

```{r}


#confusionMatrix(predict(rf_fit_overSample, songs_test_completeDF), songs_test_completeDF$counted_stream)

#confusionMatrix(predict(rf_fit_songs, songs_test), songs_test$counted_stream, positive = '1')

confusionMatrix(predict(rf_fit_underSample, songs_test), songs_test$counted_stream, positive = '1')

#confusionMatrix(predict(rf_fit_both, songs_test), songs_test$counted_stream)


```




```{r}

plot(rf_fit_underSample)

```

```{r}

importance_plot = varImpPlot(rf_fit_underSample, type = 1, scale = FALSE)

```

```{r, catch = TRUE}
library('randomForestExplainer')

plot_min_depth_distribution(rf_fit_underSample)


```


```{r}

count(songs_clean, mode)


bar_data <- songs_clean %>% group_by(mode) %>% 
  count(counted_stream)

print(bar_data)



ggplot(songs_clean, aes(x = counted_stream, y = danceability, fill = counted_stream)) + geom_boxplot()



ggplot(songs_clean, aes(x = tempo, y = us_popularity_estimate)) + geom_point() + facet_wrap(~mode) + theme_minimal()

ggplot(bar_data, aes(x = mode, y = n, fill = mode)) + geom_bar(stat = "identity") + facet_wrap(~counted_stream) + theme_bw()



```

## resampling to combat class imbalance
```{r}
library('caret')

confusionMatrix(predict(rf_fit_songs, songs_test_completeDF), songs_test_completeDF$counted_stream)

```



